---
title: "Artificial Intelligence"
author: "Group 10: Megan Lin, Eva Mustafic, James Powell"
date: "3/29/2021"
output:
  html_document:
    theme: cosmo
    toc: yes
    toc_float: yes
    df_print: paged
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r,include=FALSE}
library(tidyverse)
library(tidytext)
library(ggwordcloud)
library(gutenbergr)
library(textdata)
library(striprtf)
library(dplyr)
library(readr)
library(knitr)
# install.packages("leaflet")
library(leaflet)
```

## Background Information

Leaflet is an open-source JavaScript library for creating interactive maps that has been around since 2015.

It is used by websites ranging from The New York Times and The Washington Post to GitHub and Flickr, as well as GIS specialists at places like OpenStreetMap, Mapbox, and CartoDB to create insightful maps that offer more than just the basic photo of a map or Google Maps add-on. Its value lies in its customization that allows the developer to show more in-depth and specific features like popups, choropleths (shaded maps by density), and boundary lines.

Leaflet is currently at version 2.0.4.1 as of January 7, 2021. Here are all the past version release dates:

```{r echo = FALSE}
version<-data.frame(Version = c("1.0.0", "1.0.1", "1.1.0", "2.0.0", "2.0.1", "2.0.2", "2.0.3", "2.0.4.1"),
                    Release_Date=c("June 19, 2015", "Feb. 25, 2016", "Feb. 17, 2017", "Apr. 20, 2018", "June 4, 2018", "Aug. 27, 2018", "Nov. 14, 2019", "Jan. 7, 2021"))

version

```
(Cheng, 2021)

The main authors of the Leaflet package are Joe Cheng, Bhaskar Karambelkar, and Yihui Xie (CRAN).
### Features 

1. Interactive panning/zooming

2. Compose maps using arbitrary combinations of:
    * Map tiles
    * Markers
    * Polygons
    * Lines
    * Popups
    * GeoJSO
    
3. Create maps right from the R console or RStudio

4. Embed maps in knitr/R Markdown documents and Shiny apps

5. Easily render spatial objects from the sp or sf packages, or data frames with latitude/longitude columns

6. Use map bounds and mouse events to drive Shiny logic

7. Display maps in non spherical mercator projections

8. Augment map features using chosen plugins from leaflet plugins repository


### Installation
To install this R package, run this command at your R prompt:
```{r, eval=FALSE}
install.packages("leaflet")
library(leaflet)
```

## Usage 

### Create Maps {.tabset}

#### Blank
```{r}
leaflet() %>% addTiles() 
```

#### Satelite
```{r}
leaflet() %>% addProviderTiles(providers$Esri.WorldImagery)
```

#### Color
```{r}
leaflet() %>% addTiles(group = "OSM") %>% addProviderTiles(providers$Esri.WorldStreetMap)
```

#### Combined
```{r}
leaflet() %>% addProviderTiles(providers$Esri.WorldImagery, group= 'Satellite') %>% addProviderTiles(providers$Esri.WorldStreetMap, group = 'Street View') %>% addProviderTiles(providers$Stamen.TonerLite, group = 'Toner Lite') %>% addLayersControl(baseGroups = c("Satellite", "Street View", "Toner Lite"), position = "topright", options = layersControlOptions(collapsed = FALSE))
```

### Different Location {.tabset}

#### Zoom to Charlottesville
```{r}
leaflet() %>%
  addTiles() %>%
  setView(lat = 38.0293, lng = -78.4767, zoom = 12)
```

#### Pin the Rotunda
```{r}
leaflet() %>%
  addTiles() %>%
  # use the default base map which is OpenStreetMap tiles
  addMarkers(lat = 38.0356 , lng = -78.5034,
             popup = "The Rotunda")
```


### Create Shapes {.tabset}

#### Shape around UVA 
```{r}
leaflet() %>%
  addTiles() %>%
  addCircles(lng =-78.5080, lat=38.0336, weight = 1,
    radius = 700)
```

#### Shape around VT
```{r}
leaflet() %>%
  addTiles() %>%
  addRectangles(lng1 = -80.4384, lat1 = 37.2184 ,lng =-80.4234, lat=37.2284, weight = 1)
```
## Similar Packages 
```{r echo = FALSE}
map_packages <- data.frame(Name = c("tmap, ggplot2"), Description = c("static and interactive maps, tidyverse data visualization package"))
```
## Reflection 
## References 
(Cheng, Joe 2021) https://www.rdocumentation.org/packages/leaflet/versions/2.0.4.1
(Agafonkin, Vladimir 2016) http://rstudio.github.io/leaflet/
(CRAN) https://cran.r-project.org/web/packages/leaflet/index.html