---
title: "Artificial Intelligence"
author: "Group 7: Megan Lin, Eva Mustafic, James Powell"
date: "3/29/2021"
output:
  html_document:
    theme: cosmo
    toc: yes
    toc_float: yes
    df_print: paged
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r,include=FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(knitr)
# install.packages("leaflet")
library(leaflet)
```

## Background Information

Leaflet is an open-source JavaScript library for creating interactive maps that has been around since 2015.

It is used by websites ranging from The New York Times and The Washington Post to GitHub and Flickr, as well as GIS specialists at places like OpenStreetMap, Mapbox, and CartoDB to create insightful maps that offer more than just the basic photo of a map or Google Maps add-on. Its value lies in its customization that allows the developer to show more in-depth and specific features like popups, choropleths (shaded maps by density), and boundary lines (Agafonkin, Vladimir 2016).

Leaflet uses MapBox's maptiles to create a basemap -- a common method popularized by Google Maps and a standard for most interactive web maps. Next, tiles can be added on top of the basemap to create a specific type of overlay to show some type of information (e.g. colored weather rain maps, major roads, cartography maps, etc.) and combined to convey customized levels of information. The default tile that an be added is from OpenStreetMap and includes the basics like roadways, coloring by type of area (e.g. blue for water, green for parks, tan for general land), names of cities/states/countries, etc. Third party tiles can also be installed for more complex and customized overlays (CRAN).

Leaflet is currently at version 2.0.4.1 as of January 7, 2021. Here are all the past version release dates:

```{r echo = FALSE}
version<-data.frame(Version = c("1.0.0", "1.0.1", "1.1.0", "2.0.0", "2.0.1", "2.0.2", "2.0.3", "2.0.4.1"),
                    Release_Date=c("June 19, 2015", "Feb. 25, 2016", "Feb. 17, 2017", "Apr. 20, 2018", "June 4, 2018", "Aug. 27, 2018", "Nov. 14, 2019", "Jan. 7, 2021"))

version

```
(Cheng, 2021)

The main authors of the Leaflet package are Joe Cheng, Bhaskar Karambelkar, and Yihui Xie. Furthermore, the only dependencies are version 3.1.0 and above of R (CRAN).

### Features 

1. Interactive panning/zooming

2. Compose maps using arbitrary combinations of:
    * Map tiles
    * Markers
    * Polygons
    * Lines
    * Popups
    * GeoJSO
    
3. Create maps right from the R console or RStudio

4. Embed maps in knitr/R Markdown documents and Shiny apps

5. Easily render spatial objects from the sp or sf packages, or data frames with latitude/longitude columns

6. Use map bounds and mouse events to drive Shiny logic

7. Display maps in non spherical mercator projections

8. Augment map features using chosen plugins from leaflet plugins repository


### Installation
To install this R package, run this command at your R prompt:
```{r, eval=FALSE}
install.packages("leaflet")
library(leaflet)
```

```{r}

```

## Examples of Usage 




### Create Maps {.tabset}

Before delving deeply into the features that this package prides itself on, we should first outline the initial steps to getting a basic world map using the leaflet package. 

 * leaflet()
 * addTiles()
 * addProviderTiles()
 * addLayersControl()
 
#### addTiles()
Using the addTiles() function with no arguments in order to see the most basic world map that leaflet can deliver to you.
```{r}
leaflet() %>% addTiles() 
```

#### addProviderTiles()
addProviderTiles() is a tile adding function that is different from addTiles() for one major reason, and it is that the data argument for this function is indication that you are taking a tile template from another known map creator and using it. So in this example map, we're taking a pre-provided map that makes a realistic map of the earth in terms of imagery, and then again in the form of a roadmap.
```{r}
leaflet() %>% addProviderTiles(providers$Esri.WorldImagery)
```

```{r}
leaflet() %>% addTiles(group = "OSM") %>% addProviderTiles(providers$Esri.WorldStreetMap)
```

#### addLayersControl()
addLayersControl() allows us to layer multiple different tile sets so that the same leaflet() can create multiple different types of maps in one, and you can select which map to view.
```{r}
leaflet() %>% addProviderTiles(providers$Esri.WorldImagery, group= 'Satellite') %>% addProviderTiles(providers$Esri.WorldStreetMap, group = 'Street View') %>% addProviderTiles(providers$Stamen.TonerLite, group = 'Toner Lite') %>% addLayersControl(baseGroups = c("Satellite", "Street View", "Toner Lite"), position = "topright", options = layersControlOptions(collapsed = FALSE))
```


### Locations {.tabset}
A feature that absolutely necessary to know when using leaflet() is the ability to set the view to, pin, and label specific points on their map

#### Set View to Charlottesville
setView() is a leaflet function that allows you to zoom in on the map to a specific set of coordinates and set how intensly you want to zoom in as well!
```{r}
leaflet() %>%
  addTiles() %>%
  setView(lat = 38.0293, lng = -78.4767, zoom = 12)
```

#### Pinning
now that we have zoomed into the charlottesville area, we can easily see a pin if we placed one, to demonstrate the addMarkers() function, we will put a pin directly on the rotunda.
```{r}
leaflet() %>%
  addTiles() %>%
  # use the default base map which is OpenStreetMap tiles
  addMarkers(lat = 38.0356 , lng = -78.5034,
             popup = "The Rotunda")
```


### Create Shapes {.tabset}
Another interesting feature that is similar to the pin is the ability to add polygons to highlight areas. The examples that we have are just simple polygons, but there are many different polygons that you can creae over the map
#### Circle around UVA 
addCircles() simply adds a circle with its center located at the specified coordinates, and the radius is an argument.
```{r}
leaflet() %>%
  addTiles() %>%
  addCircles(lng =-78.5080, lat=38.0336, weight = 1,
    radius = 700)
```

#### Rectangle around VT
addRectangles() is a function similar to addCircles(), but it just superimposes a rectangle over the coordinates instead of a circle.
```{r}
leaflet() %>%
  addTiles() %>%
  addRectangles(lng1 = -80.4384, lat1 = 37.2184 ,lng =-80.4234, lat=37.2284, weight = 1)
```

##  Similar Packages
```{r echo = FALSE}
map_packages <- data.frame(Name = c("plot()", 
                                    "grid package", 
                                    "ggplot2 package", 
                                    "tmap package"), 
                           Description = c("R's built-in plotting function",
           "Allows low level control of static maps",
           "Static and interactive maps can be made, but is more generic for all types of graphs", 
           "Animated maps like a gif can be created; static and interactive maps can be mixed"))

map_packages
```
plot() is a basic r-function that can be used to create very basic maps for the user. It heavily relies on the sp and raster packages. Overall, this is a generic function that can be used to make maps, but isn't specialized for complex map making.

grid allows for low level static maps (e.g. pngs, jpgs) that allow for a snapshot of the data that is to be conveyed at a moment in time. It has low customizability, but can create more complex maps that plot()

ggplot2 is tidyverse's data visualization package that is made for create graphs of all types including bar graphs, scatterplots, line graphs, and maps. While it can create static and interactive maps and is rather easy to use because of its reliance on grammar of graphics, it isn't specifically made for maps so it can be a bit difficult to navigate and less customizable.

tmap will be more familiar to ggplot2 users as it also relies on grammar of graphics & separates aesthetics from data processing. It can create a mix of static and interactive maps and has some cool features like creating gifs and faceting graphs. This is the best contender/substitute for leaflet.

(Lovelace, Robin)

## Reflection 

## References 

(Cheng, Joe 2021) https://www.rdocumentation.org/packages/leaflet/versions/2.0.4.1
(Agafonkin, Vladimir 2016) http://rstudio.github.io/leaflet/
(CRAN) https://cran.r-project.org/web/packages/leaflet/index.html
(Lovelace, Robin)https://geocompr.robinlovelace.net/adv-map.html