---
title: "Artificial Intelligence"
author: "Group 7: Megan Lin, Eva Mustafic, James Powell"
date: "3/29/2021"
output:
  html_document:
    theme: cosmo
    toc: yes
    toc_float: yes
    df_print: paged
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r,include=FALSE}
library(tidyverse)
library(tidytext)
library(ggwordcloud)
library(gutenbergr)
library(textdata)
library(striprtf)
library(dplyr)
library(readr)
library(knitr)
# install.packages("leaflet")
library(leaflet)
```

## Background Information

Leaflet is an open-source JavaScript library for creating interactive maps that has been around since 2015.

It is used by websites ranging from The New York Times and The Washington Post to GitHub and Flickr, as well as GIS specialists at places like OpenStreetMap, Mapbox, and CartoDB to create insightful maps that offer more than just the basic photo of a map or Google Maps add-on. Its value lies in its customization that allows the developer to show more in-depth and specific features like popups, choropleths (shaded maps by density), and boundary lines.

Leaflet is currently at version 2.0.4.1 as of January 7, 2021. Here are all the past version release dates:

```{r echo = FALSE}
version<-data.frame(Version = c("1.0.0", "1.0.1", "1.1.0", "2.0.0", "2.0.1", "2.0.2", "2.0.3", "2.0.4.1"),
                    Release_Date=c("June 19, 2015", "Feb. 25, 2016", "Feb. 17, 2017", "Apr. 20, 2018", "June 4, 2018", "Aug. 27, 2018", "Nov. 14, 2019", "Jan. 7, 2021"))

version

```
(Cheng, 2021)

The main authors of the Leaflet package are Joe Cheng, Bhaskar Karambelkar, and Yihui Xie (CRAN).

### Features 

1. Interactive panning/zooming

2. Compose maps using arbitrary combinations of:
    * Map tiles
    * Markers
    * Polygons
    * Lines
    * Popups
    * GeoJSO
    
3. Create maps right from the R console or RStudio

4. Embed maps in knitr/R Markdown documents and Shiny apps

5. Easily render spatial objects from the sp or sf packages, or data frames with latitude/longitude columns

6. Use map bounds and mouse events to drive Shiny logic

7. Display maps in non spherical mercator projections

8. Augment map features using chosen plugins from leaflet plugins repository


### Installation
To install this R package, run this command at your R prompt:
```{r, eval=FALSE}
install.packages("leaflet")
library(leaflet)
```

```{r}

```

## Usage 




### Create Maps {.tabset}

Before delving deeply into the features that this package prides itself on, we should first outline the initial steps to getting a basic world map using the leaflet package. The following functions can be used to give you an initial map to work with before adding any flashy details

 * leaflet()
 * addTiles()
 * addProviderTiles()
 * addLayersControl()
 
#### addTiles()
Using the addTiles() function with no arguments in order to see the most basic world map that leaflet can deliver to you.
```{r}
leaflet() %>% addTiles() 
```

#### addProviderTiles()
addProviderTiles() is a tile adding function that is different from addTiles() for one major reason, and it is that the data argument for this function is indication that you are taking a tile template from another known map creator and using it. So in this example map, we're taking a pre-provided map that makes a realistic map of the earth in terms of imagery, and then again in the form of a roadmap.
```{r}
leaflet() %>% addProviderTiles(providers$Esri.WorldImagery)
```

```{r}
leaflet() %>% addTiles(group = "OSM") %>% addProviderTiles(providers$Esri.WorldStreetMap)
```

#### addLayersControl()
addLayersControl() allows us to layer multiple different tile sets so that the same leaflet() can create multiple different types of maps in one, and you can select which map to view.
```{r}
leaflet() %>% addProviderTiles(providers$Esri.WorldImagery, group= 'Satellite') %>% addProviderTiles(providers$Esri.WorldStreetMap, group = 'Street View') %>% addProviderTiles(providers$Stamen.TonerLite, group = 'Toner Lite') %>% addLayersControl(baseGroups = c("Satellite", "Street View", "Toner Lite"), position = "topright", options = layersControlOptions(collapsed = FALSE))
```


### Locations {.tabset}
Another feature that leaflet() includes is the ability to zoom in on, pin, and label specific points on their map

#### Zoom to Charlottesville
setView() is a leaflet function that allows you to zoom in on the map to a specific set of coordinates and set how intensly you want to zoom in as well!
```{r}
leaflet() %>%
  addTiles() %>%
  setView(lat = 38.0293, lng = -78.4767, zoom = 12)
```

#### Pinning
now that we have zoomed into the charlottesville area, we can easily see a pin if we placed one, to demonstrate the addMarkers() function, we will put a pin directly on the rotunda
```{r}
leaflet() %>%
  addTiles() %>%
  # use the default base map which is OpenStreetMap tiles
  addMarkers(lat = 38.0356 , lng = -78.5034,
             popup = "The Rotunda")
```


### Create Shapes {.tabset}

#### Circle around UVA 
```{r}
leaflet() %>%
  addTiles() %>%
  addCircles(lng =-78.5080, lat=38.0336, weight = 1,
    radius = 700)
```

#### Rectangle around VT
```{r}
leaflet() %>%
  addTiles() %>%
  addRectangles(lng1 = -80.4384, lat1 = 37.2184 ,lng =-80.4234, lat=37.2284, weight = 1)
```
## Similar Packages 
```{r echo = FALSE}
map_packages <- data.frame(Name = c("tmap, ggplot2"), Description = c("static and interactive maps, tidyverse data visualization package"))
```
(Lovelace, Robin)
## Reflection 
## References 
(Cheng, Joe 2021) https://www.rdocumentation.org/packages/leaflet/versions/2.0.4.1
(Agafonkin, Vladimir 2016) http://rstudio.github.io/leaflet/
(CRAN) https://cran.r-project.org/web/packages/leaflet/index.html
(Lovelace, Robin)https://geocompr.robinlovelace.net/adv-map.html